Scoring Patients

### 2017-07-19 Patient Report.Rmd
...

---
title: "Scoring Patients"
author: "Sarah Killcoyne"
date: "23 May 2017"
output:
  html_document: 
    toc: yes
  word_document: default
---

```{r setup, include=FALSE}
library(ggplot2)
library(GGally)
library(plyr)
library(RColorBrewer)
library(dplyr)
library(pander)
library(ggdendro)
library(reshape2)
library(GenomicRanges)
library(gridExtra)
library(tibble)

@@

## Patient info file
patient.file = grep('/All_patient_info.txt', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))

dataset = 'Training'
patient.info = read.patient.info(patient.file, set=dataset)
head(patient.info)

@@

  for (patient.name in sum.patient.data$Patient  )  {
    message(patient.name)
    patient.plot.dir = paste(plot.dir, patient.name, sep='/')

@@

    segvals = read.table(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""),sep="\t",stringsAsFactors=F,header=T)
    dim(segvals)

    missing = c(setdiff(colnames(segvals)[-(1:5)], subset(patient.info, Patient == patient.name)$Samplename),
                      setdiff(subset(patient.info, Patient == patient.name)$Samplename, colnames(segvals)[-(1:5)]))

@@

    # ORDER segvals by endoscopy year etc
    cur.pt = subset(patient.info, Patient == patient.name)
    segvals = segvals[, c(colnames(segvals[1:5]), intersect(colnames(segvals), cur.pt$Samplename))]

@@

      patient = arrange(subset(patient.info, Patient == patient.name),  Endoscopy.Year, Pathology, Samplename)
    }
  }
save(patient.data, file=cachefile)
}
print(paste("Missing patients:", paste(setdiff(sum.patient.data$Patient, names(patient.data)), collapse=', ')))

sum.patient.data = subset(sum.patient.data, Patient %in% names(patient.data))
table(sum.patient.data$Status)

## Check that all of the samples are there
patient.info = arrange(patient.info, Status, Patient, Endoscopy.Year,Samplename)
msgs = sapply(names(patient.data), function(pt) {
  missing = which(!patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals))
  if (length(missing) > 0) {
    msg = paste(pt,": Missing samples - ", paste(patient.data[[pt]]$info$Samplename[missing], collapse=', ') , sep='')
    msg = paste(msg, "\nFrom batch: ", paste(unique(patient.data[[pt]]$info$Batch.Name[missing]), collapse=', '))
    warning(msg)
  }
})
for (n in msgs) {
  if (!is.null(n))
    print(n)
}

example.pt = subset(sum.patient.data, Status == 'P' & total.samples > 2)$Patient[5]
## Note the number of probes per patient/sample
```
\pagebreak

@@

1. Segmented values (progressor): `r example.pt` 
```{r}
head(patient.data[[example.pt]]$seg.vals)
```

```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
ggplot( melt(patient.data[[example.pt]]$seg.vals[,-(1:5)]), aes(x=value) ) + 
        facet_wrap(~variable,scales = "free_x") + 
        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + labs(title=paste(example.pt,' segmented value distribution'), x='', y='')
```

2. Row (e.g. per-region across timepoints) normalized values from the matrix above
```{r}
head(patient.data[[example.pt]]$norm.seg.vals)
```

```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
ggplot( melt(patient.data[[example.pt]]$norm.seg.vals), aes(x=value) ) + 
        facet_wrap(~variable,scales = "free_x") + 
        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title=paste(example.pt,' normalized segmented value distribution'), x='', y='')

@@

sum.patient.data = merge(sum.patient.data, gen, by.x='Patient', by.y='row.names')
sum.patient.data = arrange(sum.patient.data, Status, total.samples)
sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))

@@

  do.call('grid.arrange', list(g1,g2, top=paste(name, status, sep=" - "), ncol=2))
})

@@

```{r var_reg, echo=F, message=F, warning=F}
prog.pt = subset(sum.patient.data, Status == 'P')$Patient
np.pt = subset(sum.patient.data, Status == 'NP')$Patient

var.seg.lengths = lapply(patient.data, function(df) {
  cbind('chr'=df$variable.regions$chrom, 'seg.len'=df$variable.regions$end.pos - df$variable.regions$start.pos)
})

@@

n.var.regions = sapply(var.seg.lengths, nrow)
wilcox.test(n.var.regions[prog.pt], n.var.regions[np.pt])

@@

chr.info = read.table('hg19_chrom.txt',sep='\t', header=T)

# Segment lengths, variable regions
var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chr.info, by.x='chr', by.y='chr') )

# All segments
segment.lengths = lapply(segment.lengths, function(x) merge(x, chr.info, by.x='chr', by.y='chr') )

@@

# Variable region length by ratio of chromosome
chr.ratios = lapply(var.seg.lengths, function(x) x[['seg.len']]/x[['length']])

melted = melt(chr.ratios)
melted = merge(sum.patient.data[,c('Patient', 'Status')], melted, by.x='Patient', by.y='L1')

@@

ggplot(melted, aes(Status, value, fill=Status)) + geom_boxplot() + 
  labs(title=paste("Variable region length, p-value", ifelse(chr.ratio.tt$p.value < 0.001, "<0.001", signif(chr.ratio.tt$p.value,2))),y='Segment Length/Chr length')

@@

So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(chr.ratio.tt$p.value, 2)`.  Overall bp coverage of the sWGS is `r round(mean(genome.coverage), 2)*100`%  (P `r round(mean(genome.coverage[prog.pt]),2)`, NP `r round(mean(genome.coverage[np.pt]), 2)`).

@@

@@

```{r, hmm, echo=F, message=F, warning=F, fig.height=100, fig.width=10}
ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)

noLoc = grep(':-',ccgenes$Genome.Location)
noLocGenes = ccgenes[noLoc,]

ccgenes = ccgenes[-noLoc,]

@@

  status = subset(sum.patient.data, Patient == patient.name)$Status

@@

m$brks <- cut(m$value, include.lowest=T, right=F,ordered_result=T,
                   breaks=c(0, 1, 3, 5, 8, 10),
                   labels=c('0','1-3', '4-5', '6-8','9-10'))

## TODO: Tile the variables so that you can see the gains/losses per progressor/non
subset(m, Gene.Symbols %in% c('CDKN2A', 'TBL1XR1'))
m = transform(m, Gene.Symbols = reorder(Gene.Symbols, brks))

@@

@@

m$brks <- cut(m$value, include.lowest=T, right=F,
                   breaks=c(0, 1, 3, 5, 8, 10),
                   labels=c('0','1-3', '4-5', '6-8','9-10'))

@@

@@

pvnp = data.frame(matrix(nrow=length(subset(sum.patient.data, Status == 'P')$Patient), ncol=2, dimnames=list(subset(sum.patient.data, Status == 'P')$Patient, c('hgd','be'))))    
for (name in subset(sum.patient.data, Status == 'P')$Patient) {
  var.regions = patient.data[[name]]$variable.regions
  if ( is.null(var.regions) || nrow(var.regions) == 0 ) next

@@

tt = t.test(pvnp$hgd, pvnp$be)
pvnp = cbind(pvnp, (pvnp %>% rowwise() %>% mutate(total=hgd+be, hgd.ratio=round(hgd/(hgd+be),3) ))[c('total','hgd.ratio')])

#fisher.test(p, simulate.p.value = T)

@@

pander(pvnp, justify='left', 
       caption=paste("Number of variable regions found in HGD vs non-HGD samples of progessor patients.  There's no significant difference between them (t-test p=",signif(tt$p.value,2),').  Not all patients have HGD samples (NAs). There is also no significant inverse correlation (r= ',signif(ct$estimate, 2),') between the number/ratio of variable regions found in HGD samples vs non (Pearsons p=',signif(ct$p.value, 2),')',  sep=''))

>>>>

---
title: "Scoring Patients"
author: "Sarah Killcoyne"
date: "22 June 2017"
output:
  html_document: 
    toc: yes
  word_document: default
---

```{r setup, include=FALSE}
library(ggplot2)
library(ggbio)
library(scales)
library(GGally)
library(plyr)
library(RColorBrewer)
library(dplyr)
library(pander)
library(ggdendro)
library(reshape2)
library(GenomicRanges)
library(gridExtra)
library(tibble)

@@

## Patient info file
patient.file = grep('/All_patient_info.xlsx', data.files, value=T)
if (length(patient.file) != 1)
  stop(paste("Missing/too many patient info file(s) in", data))
demo.file = grep('/Demographics_full.xlsx', data.files, value=T)

dataset = 'All'
info = read.patient.info(patient.file, demo.file, set=dataset)
removed.samples = info$removed
patient.info = info$info

#head(patient.info)

@@

  for (patient.name in sum.patient.data$Hospital.Research.ID  )  {
    message(patient.name)
    patient.plot.dir = paste(plot.dir, patient.name, sep='/')

@@

    segvals = read.table(paste(patient.plot.dir, "/",patient.name,"_segmentedCoverage_fitted_gamma",gamma2,".txt",sep=""),sep="\t",stringsAsFactors=F,header=T)
    head(segvals)

    missing = c( setdiff(colnames(segvals)[-(1:5)], subset(patient.info, Hospital.Research.ID == patient.name)$Samplename),
                 setdiff(subset(patient.info, Hospital.Research.ID == patient.name)$Samplename, colnames(segvals)[-(1:5)]) )

    mr = with(removed.samples, which(Hospital.Research.ID == patient.name & Samplename %in% missing))
    if (length(mr) > 0) {
      missing = missing[-which(missing %in% removed.samples[mr,]$Samplename)]
      segvals = segvals[,-which(colnames(segvals) %in% removed.samples[mr,]$Samplename) ]
    }

@@

    # ORDER segvals by endoscopy year etc
    cur.pt =  arrange( subset(patient.info, Hospital.Research.ID == patient.name), Endoscopy.Year, Pathology)
    segvals = segvals[, c(colnames(segvals[1:5]), intersect(colnames(segvals), cur.pt$Samplename))]

@@

      patient = arrange(subset(patient.info, Hospital.Research.ID == patient.name),  Endoscopy.Year, Pathology, Samplename)
    }
  }
save(patient.data, file=cachefile)
}
print(paste("Missing patients:", paste(setdiff(sum.patient.data$Hospital.Research.ID, names(patient.data)), collapse=', ')))

sum.patient.data = subset(sum.patient.data, Hospital.Research.ID %in% names(patient.data))
table(sum.patient.data$Status)

## Check that all of the samples are there
patient.info = arrange(patient.info, Status, Patient, Hospital.Research.ID, Endoscopy.Year,Samplename)
msgs = sapply(names(patient.data), function(pt) {
  missing = which(!patient.data[[pt]]$info$Samplename %in% colnames(patient.data[[pt]]$seg.vals))
  if (length(missing) > 0) {
    msg = paste(pt,": Missing samples - ", paste(patient.data[[pt]]$info$Samplename[missing], collapse=', ') , sep='')
    msg = paste(msg, "\nFrom batch: ", paste(unique(patient.data[[pt]]$info$Batch.Name[missing]), collapse=', '))
    warning(msg)
  }
})
for (n in msgs) {
  if (!is.null(n))
    print(n)
}

example.pt = which(with(sum.patient.data, Status == 'P' & total.samples > 2))[5]
key =  sum.patient.data[example.pt, 'Hospital.Research.ID' ]
example.pt.id = sum.patient.data[example.pt, 'Patient']
## Note the number of probes per patient/sample
```
\pagebreak

@@

1. Segmented values (progressor): `r example.pt` 
```{r}
head(patient.data[[ key ]]$seg.vals)
```

```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
ggplot( melt(patient.data[[key]]$seg.vals[,-(1:5)]), aes(x=value) ) + 
        facet_wrap(~variable,scales = "free_x") + 
        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.03) + 
  labs(title=paste('Patient', example.pt.id,' segmented value distribution'), x='', y='')
```

2. Row (e.g. per-region across timepoints) normalized values from the matrix above
```{r}
head(patient.data[[key]]$norm.seg.vals)
```

```{r echo=F, warning=F, message=F, fig.height=10, fig.width=10}
ggplot( melt(patient.data[[key]]$norm.seg.vals), aes(x=value) ) + 
        facet_wrap(~variable,scales = "free_x") + 
        geom_histogram( fill=I("blue"),col=I("lightblue"), binwidth=0.3) + labs(title=paste('Patient', example.pt.id,'normalized segmented value distribution'), x='', y='')

@@

sum.patient.data = merge(sum.patient.data, gen, by.x='Hospital.Research.ID', by.y='row.names')
sum.patient.data = arrange(sum.patient.data, Status, total.samples)
sum.patient.data$years = with(sum.patient.data, paste(start.year, end.year, sep='-'))

@@

  do.call('grid.arrange', list(g1,g2, top=paste('Patient', name, status, sep=" - "), ncol=2))
})

@@

```{r var_reg, echo=F, message=F, warning=F}
prog.pt = subset(sum.patient.data, Status == 'P')$Hospital.Research.ID
np.pt = subset(sum.patient.data, Status == 'NP')$Hospital.Research.ID

var.seg.lengths = lapply(patient.data, function(df) {
  cbind('chr'=df$variable.regions$chrom, 'start'=df$variable.regions$start.pos, 'end'=df$variable.regions$end.pos, 'seg.len'=df$variable.regions$end.pos - df$variable.regions$start.pos)
})

@@

sum.var = rbind(
  'Progressors'=do.call(rbind.data.frame, var.seg.lengths[prog.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)),
  'Non-Progressors'=do.call(rbind.data.frame, var.seg.lengths[np.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)))

pander(sum.var, caption='Variable regions between progressors and non-progressors')

sum.seg = rbind(
  'Progressors'=do.call(rbind.data.frame, segment.lengths[prog.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)),
  'Non-Progressors'=do.call(rbind.data.frame, segment.lengths[np.pt]) %>% summarise( 'n.segments' = length(seg.len), 'median.length'=median(seg.len), 'mean.length'=mean(seg.len), 'sd'=sd(seg.len)))

pander(sum.var, caption='Segment stats between progressors and non-progressors')
```

```{r echo=F, message=F, warning=F}

df = rbind(cbind(do.call(rbind.data.frame, var.seg.lengths[prog.pt]), 'status'='Progressor'), cbind(do.call(rbind.data.frame, var.seg.lengths[np.pt]), 'status'='Non-Progressor'))
lenWT = wilcox.test(subset(df, status == 'Progressor')$seg.len, subset(df, status == 'Non-Progressor')$seg.len)

ggplot(df, aes(status, seg.len)) + geom_jitter(aes(color=status), alpha=0.5, show.legend = F, width=.2) +
  geom_boxplot(aes(group=status, fill=status), show.legend=F, outlier.colour = NA) + 
  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(seg.len)-2*sd(seg.len), label=paste('median\n',scientific(median(seg.len),3),sep='') )) +
  geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(seg.len), label=paste('median\n',scientific(median(seg.len),3),sep=''))) +
  labs(y='Variable segment length', x='', title='Variable regions', caption=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))

ggplot(df, aes(status, seg.len)) + 
  geom_jitter(aes(color=status), show.legend=F) + geom_violin(aes(group=status, fill=status), show.legend=F) +
  geom_label(data=subset(df, status == 'Progressor'), aes(status, max(seg.len)-2*sd(seg.len),label=paste('median\n',scientific( median(seg.len) , 3),sep='') ))  +
    geom_label(data=subset(df, status == 'Non-Progressor'), aes(status, max(seg.len), label=paste('median\n',scientific( median(seg.len) , 3),sep=''))) +
    labs(y='Variable segment length', x='', title='Variable regions', caption=paste('Wilcox p-value', signif(lenWT$p.value, digits=3) ))

```{r, echo=F, warning=F, message=F}
df = rbind(cbind.data.frame('n'=sapply(var.seg.lengths[prog.pt], nrow), 'status'='Progressor'), cbind.data.frame('n'=sapply(var.seg.lengths[np.pt], nrow), 'status'='Non-Progressor'))
nWT = wilcox.test(subset(df, status == 'Progressor')$n, subset(df, status == 'Non-Progressor')$n)

ggplot(df, aes(status, n, group=status)) + geom_jitter(aes(color=status), width=0.2) +
  geom_boxplot(aes(fill=status), position = 'dodge', show.legend = F, outlier.colour = NA) +
  geom_label(data=subset(df, status == 'Progressor'), 
             aes(status, max(n)-2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status ), color='white' ) +
  geom_label(data=subset(df, status == 'Non-Progressor'), 
             aes(status, max(n)+2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status), color='white') +
  labs(y='Number of segments', title="Total number of segments", caption=paste('Wilcox p-value', signif(nWT$p.value, digits=3)) ) +
  theme(legend.position = 'none')

ggplot(df, aes(status, n, group=status)) + geom_jitter(aes(color=status), width=0.2) +
  geom_violin(aes(fill=status, group=status)) +
  geom_label(data=subset(df, status == 'Progressor'), 
             aes(status, max(n)-2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status ), color='white' ) +
  geom_label(data=subset(df, status == 'Non-Progressor'), 
             aes(status, max(n)+2*sd(n), label=paste('median\n',sprintf("%.1f",median(n)),sep=''), fill=status), color='white') +
  labs(y='Number of segments', title="Total number of segments", caption=paste('Wilcox p-value', signif(nWT$p.value, digits=3)) ) +
  theme(legend.position = 'none')

### Check variable segment length vs chromosome length

```{r echo=F, message=F, warning=F, fig.height=8, fig.width=20}
n.var.regions = sapply(var.seg.lengths, nrow)

#wilcox.test(n.var.regions[prog.pt], n.var.regions[np.pt])

@@

chr.info = read.table('hg19_chrom.txt',sep='\t', header=T)
chr.info$genome.length = cumsum(as.numeric(chr.info$length))

var.seg.lengths = lapply(var.seg.lengths, function(x) merge(x, chr.info, by='chr') )

# All segments
segment.lengths = lapply(segment.lengths, function(x) merge(x, chr.info, by='chr') )

@@

var.seg.lengths = lapply(var.seg.lengths, function(x) cbind(x, 'chr.ratio'=x[['seg.len']]/x[['length']]))

# Segment lengths, variable regions
vsl = do.call(rbind.data.frame, lapply(patient.data, function(df) {
  if (is.null(df$variable.regions) || nrow(df$variable.regions) <= 0) return(NULL)

  mc = grep('^D', colnames(df$variable.regions))

  meanCN = sapply(apply(t(df$variable.regions[,mc]), 2, function(x) x[which(x >= 1.1 | x <= 0.9)]), mean)
  meanCN[is.na(meanCN)] = 0

    cbind.data.frame( 'chr'=df$variable.regions$chrom, 'start'=df$variable.regions$start.pos, 'end'=df$variable.regions$end.pos, meanCN, 'status'= df$info$Status[1])
  }))
vsl = vsl[-which(vsl$meanCN == 0),]

vsl$meanCN = vsl$meanCN - mean(vsl$meanCN)

vslm = melt(vsl, id.vars=c('chr','start','end','status'))

chr.info = subset(chr.info, chr %in% c(1:22))
chr.info$chr = ordered(chr.info$chr, levels=c(1:22))
vslm$chr =  ordered(vslm$chr, levels=c(1:22))
vslm = merge(chr.info, vslm, by='chr')
vslm = vslm %>% rowwise() %>% mutate('genome.start'=genome.length+start, 'genome.end'=genome.length+end)

vslm$status = ifelse(vslm$status == 'P', 'Progressor', 'Non Progressor')

ggplot(vslm, aes(x=length)) + 
  facet_grid(status~chr) +
  geom_rect(data=vslm, aes(xmin=start, xmax=end, ymin=0, ymax=value, fill=value)) +
  scale_fill_gradient(low = "green3", high = "red3", name='') + 
  labs(title="Variable region gains & losses", x='Chromosomes', y='Mean adjusted segmentation value') +
  theme(axis.text.x=element_blank(),axis.ticks.x=element_blank(), legend.position='bottom')

```{r echo=F, message=F, warning=F}
# Variable region length by ratio of chromosome
chr.ratios = lapply(var.seg.lengths, function(x) x[['seg.len']]/x[['length']])

melted = melt(chr.ratios)
melted = merge(sum.patient.data[,c('Hospital.Research.ID', 'Status')], melted, by.x='Hospital.Research.ID', by.y='L1')

@@

ggplot(melted, aes(Status, value, fill=Status)) + geom_jitter(aes(color=Status), width=0.2) + geom_boxplot(outlier.colour = NA) + 
  geom_label(data=subset(melted, Status == 'P'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white', show.legend = F ) +
  geom_label(data=subset(melted, Status == 'NP'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
  labs(title="Variable region length",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Segment Length/Chr length')

ggplot(melted, aes(Status, value, fill=Status)) + geom_jitter(aes(color=Status), width=0.2) + geom_violin() + 
  geom_label(data=subset(melted, Status == 'P'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
  geom_label(data=subset(melted, Status == 'NP'), aes(x=Status, y=max(value)-sd(value), label=scientific(median(value),3) ), color='white' , show.legend = F) +
  labs(title="Variable region length",subtitle=paste("p-value", signif(chr.ratio.tt$p.value, digits=3)), y='Segment Length/Chr length')

@@

So it appears that adjusting for chromosome length (looking at the ratio of the chromosome covered by segments rather than the raw bp) a significant difference in the coverage of the genome by CN segments that are variable between progressors and non: t-test p-value `r signif(chr.ratio.tt$p.value, 3)`.  Overall bp coverage of the sWGS is `r round(mean(genome.coverage), 2)*100`%  (P `r round(mean(genome.coverage[prog.pt]),2)`, NP `r round(mean(genome.coverage[np.pt]), 2)`).

@@

## Gain vs Loss

```{r echo=F, message=F, warning=F}

df = do.call(rbind.data.frame, lapply(patient.data, function(df) {
  nohgd = subset(df$info, !Pathology %in% c('HGD','IMC'))$Samplename
  gain=0;loss=0
  if ( !is.null(df[['variable.regions']]) ) {
    nohgd = intersect(nohgd, colnames(df$variable.regions))
    gain = length(which(apply(as.data.frame(df$variable.regions[,nohgd]), 1, function(x) length(which(x >= 1.1))) > 0))
    loss = length(which(apply(as.data.frame(df$variable.regions[,nohgd]), 1, function(x) length(which(x <= 0.9))) > 0))
  }

  cbind.data.frame('gain'=gain, 'loss'=loss, 'status'=unique(df$info$Status))
}))
head(df)

ttg = t.test(subset(df, status == 'P')$gain, subset(df, status == 'NP')$gain)
ttl = t.test(subset(df, status == 'P')$loss, subset(df, status == 'NP')$loss)

ft = fisher.test((df %>% group_by(status) %>% summarise( 'gains'=sum(gain), 'losses'=sum(loss)))[,c('gains','losses')])

ggplot(melt(df), aes(status, value, fill=status)) + facet_grid(.~variable) + geom_bar(stat='identity') +
  labs(title='Total gains & Losses excluding HGD/IMC', y='counts', x='', caption=paste('Gains and losses are signficantly enriched in Progressors vs Non.\nFisher odds ', round(ft$estimate,2), '(p=', signif(ft$p.value, digits=3), ')', sep='' ))

@@

```{r, hmm, echo=F, message=F, warning=F, fig.height=60, fig.width=10}
ccgenes = read.table('~/Data/CosmicCensusGenes.tsv', sep='\t', header=T, stringsAsFactors=F)
ccgenes = ccgenes[-which(ccgenes$Mutation.Types == 'T'),]

noLoc = grep(':-',ccgenes$Genome.Location)
if (length(noLoc) > 0) {
  noLocGenes = ccgenes[noLoc,]
  ccgenes = ccgenes[-noLoc,]
}

@@

  status = subset(sum.patient.data, Hospital.Research.ID == patient.name)$Status

@@

m$brks <- cut(m$value, include.lowest=T, right=F,ordered_result=T,
                   breaks=c(0, 1, 3, 5, 8, 10, 12, 15),
                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12', '>12'))

## TODO: Tile the variables so that you can see the gains/losses per progressor/non
#subset(m, Gene.Symbols %in% c('CDKN2A', 'TBL1XR1'))
m = transform(m, Gene.Symbols = reorder(Gene.Symbols, brks))

@@

m2 = melt(gene.p %>% group_by(Gene.Symbols) %>% summarise( 'Progressors' = sum(Progressors), 'Non.Progressors' = sum(Non.Progressors)), id.vars='Gene.Symbols')

m2 = transform(m2, Gene.Symbols = reorder(Gene.Symbols, value))

m2$brks <- cut(m2$value, include.lowest=T, right=F,ordered_result=T,
                   breaks=c(0, 1, 3, 5, 8, 10, 12, 15),
                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12', '>12'))

ggplot(m2,  aes(y=Gene.Symbols, x=variable, fill=brks)) + geom_tile() + #facet_grid(.~variable) + 
    scale_fill_manual(values = myP, 'Patient Counts') + labs(title="All available COSMIC genes", x='', y='')

@@

m$brks <- cut(m$value, include.lowest=T, right=F,
                   breaks=c(0, 1, 3, 5, 8, 10, 12, 15),
                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12', '>12'))

@@

gene.p50 = gene.p %>% group_by(Gene.Symbols) %>% summarise( 'Progressors' = sum(Progressors), 'Non.Progressors' = sum(Non.Progressors))

m2 = melt(subset(gene.p50, Gene.Symbols %in% gc[1:50, 'Gene.Symbols']), id.vars=c('Gene.Symbols'))

m2 = transform(m2, Gene.Symbols = reorder(Gene.Symbols, value))

m2$brks <- cut(m2$value, include.lowest=T, right=F, ordered_result=T,
                   breaks=c(0, 1, 3, 5, 8, 10, 12, 15),
                   labels=c('0','1-3', '4-5', '6-8','9-10', '10-12', '>12'))

ggplot(m2,  aes(y=Gene.Symbols, x=variable, fill=brks)) + geom_tile(color='white') + 
    scale_fill_manual(values=myP, 'Patient Counts') + 
    labs(title="Top 50 genes by number of patients", x='', y='')

@@

pvnp = data.frame(matrix(nrow=length(subset(sum.patient.data, Status == 'P')$Hospital.Research.ID), ncol=2, dimnames=list(subset(sum.patient.data, Status == 'P')$Hospital.Research.ID, c('hgd','be'))))    
for (name in subset(sum.patient.data, Status == 'P')$Hospital.Research.ID) {
  var.regions = patient.data[[name]]$variable.regions
  if ( is.null(var.regions) || nrow(var.regions) == 0 ) next

@@

#tt = t.test(pvnp$hgd, pvnp$be)
pvnp = cbind(pvnp, (pvnp %>% rowwise() %>% mutate(total=hgd+be, hgd.ratio=round(hgd/(hgd+be),3) ))[c('total','hgd.ratio')])

ft = fisher.test( pvnp[complete.cases(pvnp[,c('hgd','be')]), c('hgd','be')], simulate.p.value = T)

@@

pander(pvnp, justify='left', 
       caption=paste("Number of variable regions found in HGD vs non-HGD samples of progessor patients.  There's a significant difference between them (Fisher's p=",signif(ft$p.value,3),').  Not all patients have HGD samples (NAs). There is no significant correlation (r= ',signif(ct$estimate, 2),') between the number/ratio of variable regions found in HGD samples vs non (Pearsons p=',signif(ct$p.value, 3),')',  sep=''))
